---
title: "wranglingGapminder"
author: "Jasper Wong"
date: "February 22, 2017"
output: github_document
---

##Introduction

This analysis profiles a look into the [gapminder](https://github.com/jennybc/gapminder) dataset, as per the course [STAT545](http://stat545.com/) taught by Jenny Bryan.

Gapminder is a well known dataset that highlights a number of descriptors for countries, including Gdp per capita, life expectancy, and population throughout years of data collection.

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressWarnings(suppressPackageStartupMessages(library(ggplot2)))
suppressWarnings(library(reshape2))
suppressWarnings(suppressPackageStartupMessages(library(tidyverse)))
suppressWarnings(library(knitr))
```

```{r, echo = FALSE}
gapminder <- read.delim(file = "http://www.stat.ubc.ca/~jenny/notOcto/STAT545A/examples/gapminder/data/gapminderDataFiveYear.txt")
```

##Basic analyses

###Gdp per capita:

Looking at how gdp per capita changes over time.
```{r, echo = FALSE, results = 'hide'}

p <- ggplot(gapminder, aes(x = gdpPercap,  y = lifeExp, colour = continent, size = pop)) + geom_point() + scale_x_log10()

p + facet_wrap(~year)

```


A heatmap can be generated as follows.

```{r, echo = FALSE, message = FALSE, results = 'hide'}
#generating a heat map of gdp per cap's change over the years for every country

library(pheatmap)
library(RColorBrewer)
matrixcolors <- brewer.pal(8, "RdBu")
gapminder$log_gdpPercap <- log10(gapminder$gdpPercap)
gapminder_gdp_years <- subset(gapminder, select = c(country, year, log_gdpPercap))
gdpYears <- spread(gapminder_gdp_years, year, log_gdpPercap)
gdpYears_matrix <- as.matrix(sapply(gdpYears[,2:13], as.numeric))
rownames(gdpYears_matrix) <- gdpYears[,1]
pheatmap(gdpYears_matrix, color = matrixcolors, cluster_cols = FALSE)
```

